Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-12-20 write_up/reviewer_response.tex

\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{booktabs}

\begin{document}

\large

Address here: cover letter on other computer. \\
\\
Dear xyz, \\
RE: Ms. No. ECOMOD-19-631 \\
\\
Thank you for your email of 5 December 2019 with the favourable 'minor
revisions` review of our manuscript submission. We have reviewed all the
comments made by the editor and the two reviewers and addressed them
accordingly. We thank the reviewers for a constructive review which undoubtedly
improves the manuscript and we found little reason to disagree with their
comments. A summary of our changes are presented overleaf.  \\

@@

\\
Why was the Matern covariance structure chosen? I am just curious about this;
does it reflect the spatial autocorrelation of animal distributions better than
a Gaussian or other type of covariance structure? & We used the Matérn
covariance structure as it allows for a non-linear decrease in correlation as
distance increases and is flexible, in that it reverts to an exponential form
under specific conditions. As such it is capable of taking any number of forms,
as required. Alternative covariance structures could also be used, but Matérn
is the most commonly used. We've added text to say "We use the most commonly
used Matérn covariance structure as it is a flexible form that under certain
conditions is of the same form as an exponential function and it enables us to
model the spatial autocorrelation observed in animal populations where density
is more similar in nearby locations, but that correlation decreases
non-linearly...." \\ 

@@

\\
It would be helpful to have explanation of thermal tolerance earlier than line
155 & We have now added a sentence at the beginning of the section that reads
"[described by a set of probabilities] which are affected by the suitability of
habitat, temperature in a cell and the thermal tolerance of a population to
that temperature." \\

@@

\\
The paragraph starting at line 179 would be helpful to have at the beginning of
the section & Agree, we have moved this to now be the second paragraph in the
section. \\

@@

\\
Line 187, make tenses consistent (e.g. determined vs. determines) - also this
sentence is a little awkward to read. & We have changed so that it reads in the
past tense and rewritten the paragraph to improve readability. \\

### 2019-12-20 write_up/mixfishsim_ecolmod_revisions.tex
Updated response to reviewers

Population movement is a combination of directed (advective) movement where at
certain times of year the population moves towards spawning grounds by
increasing the probabilities of moving into the spawning grounds from adjacent
cells, and random (diffusive) movement, governed by a stochastic process where
movement between adjacent cells is described by a set of probabilities.\\

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. This enables us to model the spatial
autocorrelation observed in animal populations where density is more similar in
nearby locations \citep{Tobler1970, F.Dormann2007, Poos2007} and we change the
parameters to implement different spatial structures for the different
populations using the \textit{RandomFields} R package \citep{Schlater2015}. We
define a stationary habitat field with an anisotropic pattern (to simulate a
depth gradient) and combine it with a temporally dynamic thermal tolerance
field to imitate two key drivers of population dynamics without modelling the
processes explicitly. Each population was initialised at a single location, and
subsequently moved across the entire space according to a probabilistic
distribution based on habitat suitability (represented by the normalised values
from the GRFs), temperature tolerance and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}$ is the index of habitat
suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see below).
\\

@@

The final combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).

@@

Fleet dynamics can be broadly categorised into three components: fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular population; trip-level decisions, that determines the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determines movement from one fishing spot to another within a trip. An
explore-exploit type strategy was implemented in the model that combined these
three components for individual vessels to maximise their catch from an unknown
resource distribution \citep{Bailey2018}. The decision to use an individual
based model for fishing vessels was taken because fishers are heterogeneous in
their location choice behaviour due to different objectives, risk preference
and targeting preference \citep{VanPutten2012a, Boonstra2015}. Therefore fleet
dynamics are emergent from individual dynamics rather than pre-defined group
dynamics.

@@

\section{Parameterisation}
We parameterise \textit{MixFishSim} to investigate the influence of data
aggregation on spatial inference.

@@

We parametrised the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
parameters (Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature (Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4})
defined to be unique to each population resulting in differently weekly
distribution patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each
of the populations was assumed to have two defined spawning areas that result
in the populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}).

\subsection{Fleet parametrisation}

The fleets were parametrised to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}). By setting different
catchability parameters ($Q_{fl, p}$) we create different targeting preferences
between the fleets and hence spatial dynamics.  The learned random walk process
implies that within a fleet different vessels have different spatial
distributions based on individual experience. The step function was
parametrised dynamically within the simulations as the maximum revenue
obtainable was not known beforehand. This was implemented so that vessels take
smaller steps when fishing at a location that yields landings value in the top
90th percentile of the value experienced in that year so far (as defined per
fleet in Table \ref{tab:2}). \\

@@

The aggregated catch composition from each of the data sources over a ten-year
period (which shows average seasonal patterns) at different spatial resolutions
highlights different patterns in perceived community structure depending on the
data source and aggregation level (Figure \ref{fig:1}; visualised using
\cite{Gerritsen2014}). The finer spatial grid for the true population (top
left) and commercial data (top middle) show visually similar patterns, though
there are large unsampled areas in the commercial data from a lack of fishing
activity (particularly in the lower left part of the sampling domain). The
survey data at this spatial resolution displays very sparse information about
the spatial distributions of the populations. The slightly aggregated data on a
5 x 5 grid shows similar patterns and, while losing some of the spatial detail,
there remains good consistency between the true population and the commercial
data. Survey data starts to pick out some of the similar patterns as the other
data sources, but
lacks spatiotemporal coverage. The spatial catch information on a 10 x 10 and
20 x 20 grid lose a significant amount of information about the spatial
resolutions for all data sources, and some differences between the survey,
commercial and true population data emerge. \\

@@

Fish populations in our simulations move in pre-defined timescales and
according to fixed habitat preferences and temperature gradients (Figures
\ref{S-fig:1}, \ref{S-fig:3}). Our assumptions in parameterising the model
(movement rates, temperature tolerances) will have a direct impact on our
conclusions on the relative importance of spatial and temporal processes. These
assumptions could be explored in a future study by varying the parameters and
assessing the robustness of our conclusions. For our example application we
have chosen movement rates to reflect aggregation periods observed in past
studies \citep{Poos2007}. \\

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{[Colour] Data aggregation at different spatial resolutions over a
		ten year period.}
	\label{fig:1}
\end{figure}

>>>>

Population movement is a combination of directed (advective) movement where at
certain times of year the population moves towards spawning grounds by
increasing the probabilities of moving into the spawning grounds from adjacent
cells, and random (diffusive) movement, governed by a stochastic process where
movement between adjacent cells is described by a set of probabilities
\added{which are affected by the suitability of habitat, temperature in a cell
	and the thermal tolerance of a population to that temperature}.\\

\added{The combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns }(Figure \ref{S-fig:5}).

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. \added{We use the most commonly used
	Matérn covariance structure as it is a flexible form that under certain
	conditions is of the same form as an exponential function and
	it}\deleted{This} enables us to model the spatial autocorrelation
observed in animal populations where density is more similar in nearby
locations, \added{but that correlation decreases non-linearly}
\citep{Tobler1970, F.Dormann2007, Poos2007}. We change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field with an anisotropic pattern (to simulate a depth gradient) and
combine it with a temporally dynamic thermal tolerance field to imitate two key
drivers of population dynamics without modelling the processes explicitly. Each
population was initialised at a single location, and subsequently moved across
the entire space according to a probabilistic distribution based on habitat
suitability (represented by the normalised values from the GRFs), temperature
tolerance and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}$ is the index of habitat
suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see below).
\\

@@

\deleted{The final combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).}

@@

Fleet dynamics \replaced{were}{can be }broadly categorised into three
components. \textit{Fleet targeting} \replaced{determined}{determines} the
fleet catch efficiency and preference towards a particular population;
\textit{trip-level decisions} \replaced{determined}{that determines} the
initial location to be fished at the beginning of a trip; and
\textit{within-trip decisions}\replaced{determined fishing locations}{, that
	determines movement from one fishing spot to another} within a trip.
\replaced{This results in an explore-exploit strategy}{ was implemented in the
	model that combined these three components} for individual vessels to
maximise their catch from an unknown resource distribution \citep{Bailey2018}.
The decision to use an individual based model for fishing vessels was taken
because fishers are heterogeneous in their location choice behaviour due to
different objectives, risk preference and targeting preference
\citep{VanPutten2012a, Boonstra2015}.  Therefore fleet dynamics are emergent
from individual dynamics rather than pre-defined group dynamics.

@@

\section{Model calibration}
We calibrate \textit{MixFishSim} to investigate the influence of data
aggregation on spatial inference.

@@

We calibrated the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
(Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature (Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4})
defined to be unique to each population resulting in differently weekly
distribution patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each
of the populations was assumed to have two defined spawning areas that result
in the populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}).

\subsection{Fleet calibration}

The fleets were calibrated to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}). By setting different
catchability coefficients ($Q_{fl, p}$) we create different targeting
preferences between the fleets and hence spatial dynamics.  The learned random
walk process implies that within a fleet different vessels have different
spatial distributions based on individual experience. The step function was
calibrated dynamically within the simulations as the maximum revenue
obtainable was not known beforehand. This was implemented so that vessels take
smaller steps when fishing at a location that yields landings value in the top
90th percentile of the value experienced in that year so far (as defined per
fleet in Table \ref{tab:2}). \\

@@

The \deleted{aggregated }catch composition\added{aggregated at different
	spatial resolutions} from each of the data sources (\deleted{which
	shows }average seasonal patterns over a ten-year period) highlights
different patterns in perceived community structure depending on the data
source and aggregation level (Figure \ref{fig:1}). The finer spatial grid for
the true population (top left) and commercial data (top middle) show visually
similar patterns, though there are large unsampled areas in the commercial data
from a lack of fishing activity (particularly in the lower left part of the
sampling domain). The survey data at this spatial resolution displays very
sparse information about the spatial distributions of the populations. The
slightly aggregated data on a 5 x 5 grid shows similar patterns and, while
losing some of the spatial detail, there remains good consistency between the
true population and the commercial data. Survey data starts to pick out some of
the similar patterns as the other data sources, but lacks spatiotemporal
coverage. The spatial catch information on a 10 x 10 and 20 x 20 grid lose a
significant amount of information about the spatial resolutions for all data
sources, and some differences between the survey, commercial and true
population data emerge. \\

@@

Fish populations in our simulations move in pre-defined timescales and
according to fixed habitat preferences and temperature gradients (Figures
\ref{S-fig:1}, \ref{S-fig:3}). Our assumptions in calibrating the model
(movement rates, temperature tolerances) will have a direct impact on our
conclusions on the relative importance of spatial and temporal processes. These
assumptions could be explored in a future study by varying the parameters and
assessing the robustness of our conclusions. For our example application we
have chosen movement rates to reflect aggregation periods observed in past
studies \citep{Poos2007}. \\

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{[Colour] Data aggregation at different spatial resolutions over a
		ten year period. The figure shows catch composition at each
		spatial unit as visualised using \cite{Gerritsen2014} mapplots
		R package.}
	\label{fig:1}
\end{figure}
